# Welfare

In WITCH, a social planner maximises the total discounted welfare $W$ of each coalition $clt$.
The regional utility function at any point in time and each region is based on a power or 
constant relative risk aversion (CRRA) function of the consumption per capita. 
Given the type of cooperation, the objective function can take different forms.

## Utility function

For each region $n$, intertemporal utility is computed as discounted sum of utility: 

$$ W(n) = \sum_{t} l(t,n) \frac{\left(\frac{C(t,n)}{l(t,n)}\right)^{1-\eta}-1}{1-\eta} \beta^t. $$

The pure time preference discount factor $\beta$ in any case is given by the geometric discounting rule:

$$ \beta = ({1+\rho})^{\Delta_{t}} $$

where $\Delta_{t}=5$ is the duration of one time step in years.


## Welfare of coalitions

The WITCH model can be run in a non-cooperative mode, where each region $n$ acts as one player maximising its welfare. In this case, the set of players or coalitions $clt$ consists each of a single region.

In the cooperative mode, different coalitions can be formed including the Grand Coalition containing all regions.
With cooperation, there are hence one or several coalitions $clt$, who maximise welfare in their member region(s). The coalition $clt$ maximises a sum of the welfare of the member regions. There are several welfare concepts admissible to aggregate welfare across coalition members. 

The default option is a (disentangled) Utilitarian solution, taking into account inequality across regions through a degree of inequality aversion. This welfare concept is related to Epstein-Zin preferences and welfare of a coalition $clt$ is given by the equation

$$ W(clt)= \sum_{t=1}^{T} \left[\frac{1}{1-\eta}\left(\left(\sum_{n \in clt}l(t,n) (\frac{C(t,n)}{l(t,n)} \right )^{1-\gamma}\right)^{\frac{1-\eta}{1-\gamma}}-1\right] \beta^t $$

Note that by default, we consider no inequality aversion($gamma=0$), since this equalises marginal welfare across regions, leads to a unique social cost of carbon, and is quantitatively similar to the common use of Negishi weights [@Negishi1960]. As an alternative, time varying Negishi weights [@NordhausYang1996] can be used in which case the welfare function has the form of

$$ W^{Negishi}(clt) = \sum_{t}\sum_{n\in clt} w_{t,n} l(t,n) \frac{\left(\frac{C(t,n)}{l(t,n)}\right)^{1-\eta}-1}{1-\eta} \beta^t$$

where $w_{t,n}$ are the time- and region-specific Negishi weights which are computed as 

$$ w_{t,n}=\frac{\frac{1}{c(t,n)^{\eta}}}{\sum_{n' \in clt}\frac{1}{c(t,n')^{-\eta}}}. $$

```{r, echo=F}
library(data.table)
sym_witch = fread("symbols.csv",sep=",",header=T)
tab = sym_witch[id %in% 1:5,.(Symbol,Definition,GAMS,Unit)]
knitr::kable(
tab, booktabs = TRUE,
caption = 'Welfare symbolic terms'
)
```
	
For the current version of WITCH, the parameters are reproduced in Table \@ref(tab:welfparam).

```{r welfparam, echo=F, tidy=FALSE}
library(data.table)
sym_witch = fread("symbols.csv",sep=",",header=T)
tab = sym_witch[Symbol %in% c("$\\eta$","$\\rho$","$\\gamma$"),.(Symbol,Definition,GAMS)]
tab[,`Default value`:=c(1.5,0.01,0)]
knitr::kable(
tab, booktabs = TRUE,
caption = 'Welfare parameters'
)
```